Less Funding, More Lecturers, and Fewer Professors: \\ \vspace{0.1cm} Stagnating State Funding for Higher Education and its Effect on Faculty

### 2024-05-23 text/sections/05-results.tex
Small fixes.

To investigate reduced hiring of new professors, I first investigate faculty hiring rates among Illinois universities from the IBHED data.
To do so, I count the number of faculty new in the panel of all Illinois public university faculty for years 2011--2021, where the panel first year is 2010.
I then use this as the outcome in OLS and shift-share IV regressions, where each observation is an Illinois university--year and data for state funding come from IPEDS.\footnote{
    This is the same specification as the IPEDS analysis restricted to Illinois universities.
    The IPEDS data have yearly faculty counts, but not discern between faculty promotions, quits, and hires; the Illinois data can do so thanks to the panel structure of the data.
}
To study whether these effects hold true for nation-wide hiring of new professors, I use public on the number of professors at all US public universities for the entire decade 2011--2021, and perform the same analysis to see if state funding affected the number of professors hired.

\autoref{tab:facultyhires-illinois-reg} shows these estimates for Illinois public universities.
In general, faculty counts are relatively unrelated to state funding among Illinois universities; though the point estimates line up with analysis in IPEDS, the smaller number of observations mean that the precision of estimates are small and indistinguishable from zero, in bith raw counts and in log units.
On the other hand, universities with more state funding hire more professors across 2011--2021, as seen in \autoref{fig:hiring-correlation}.
An extra 10\% in state funding is associated with 8.5\% more professors hired per students.
I use these data as part of a shift-share analysis, showing that a 10\% state funding cut per student leads to 13\% fewer professor hires (relative to enrolment), and that these hires are roughly equal for gender of professor hired.

This shows that state funding is related in to professor hiring, and that the persistent funding cuts to public universities led to lower professor hiring across this time period.
The other turn-over channels were unaffected by state funding cuts, and that professor hiring is related to professor hiring in both correlational terms and in a shift-share analysis.
This means that effects on faculty hiring were likely the main mechanism for how state funding cuts led to a substitution of professors for lecturers at public universities.

>>>>

To investigate reduced hiring of new professors, I first investigate faculty hiring rates among Illinois universities from the IBHED data.
To do so, I count the number of faculty new in the panel of all Illinois public university faculty for years 2011--2021, where the panel first year is 2010.
I then use this as the outcome in OLS and shift-share IV regressions, where each observation is an Illinois university--year and data for state funding come from IPEDS.\footnote{
    This is the same specification as the IPEDS analysis restricted to Illinois universities.
    The IPEDS data have yearly faculty counts, but not discern between faculty promotions, quits, and hires; the Illinois data can do so thanks to the panel structure of the data.
}
To study whether these effects hold true nation-wide, I use public on the number of professors at all US public universities for the entire decade 2011--2021, provided by \citep{wapman2022quantifying},
and perform the same analysis to see if state funding affected the number of professors hired.
See \autoref{sec:appendix-hiring} for further details on this data source.

\autoref{tab:facultyhires-illinois-reg} shows these estimates for Illinois public universities.
In general, faculty counts are relatively unrelated to state funding among Illinois universities; though the point estimates line up with analysis in IPEDS, the smaller number of observations mean that the precision of estimates are small and indistinguishable from zero, in both raw counts and in log units.
On the other hand, universities with more state funding hire more professors across 2011--2021, as seen in \autoref{fig:hiring-correlation}.
An extra 10\% in state funding is associated with 8.5\% more professors hired per students.
I use these data as part of a shift-share analysis, showing that a 10\% state funding cut per student leads to 13\% fewer professor hires (relative to enrolment), and that these hires are roughly equal for gender of professor hired.

State funding cuts lead to lower professor hiring, and the persistent funding cuts to public universities led to lower professor hiring across the last three decades.
The channels for faculty turn-over were unaffected by state funding cuts in the analysis of Illinois faculty, and rate professor hiring is related to professor hiring in both correlational terms and in a shift-share analysis.
This means that effects on faculty hiring were likely the main mechanism for how state funding cuts led to a substitution of professors for lecturers at public universities.
